Toward theories of deliberated preferences

### 2020-12-07 simple.tex
Attempt fals using claim of Convincingness, interrupted as well because difficult to describe convincingness.

\subsection{Validity and falsification sequences}
\label{sec:validity}
An empirical theory claims to be valid, in the following sense.
\begin{definition}[Validity]
	An empirical theory is \emph{valid} iff $\forall i \in I, \prop \in \gpropse: \prop \in \iprops$.
\end{definition}
To achieve this goal, it claims that it is able to produce decisive arguments for its supported propositions.
These arguments can vary according to the individual $i \in I$.
Given an individual, the theory claims that a decisive argument will be systematically found by starting from an initial argument given by $\gleadsto$, and following a chain of defense defined by $\dg$ as required, depending on the paths that challenge the arguments produced along the way, as judged by the individual $i$. The property that a theory manages to systematically produce decisive arguments in such a way is named Convincingness. Defining it requires to define first the notion of required arguments.

@@

To verify that the property of Convincingness is purely empirical, it is necessary to show that whenever a theory fails it, it is possible to see it using a set of observations of $\allleadsto$. Conversely, it is also desirable to ensure that only the observations that ensure that Convincingness is false count for falsifying a theory.
The following example may help to form an intuition about what the possible falsifying sets of observations are.
\begin{example}[Falsification sequences]
	Let us examine the possible observations that would suffice to show that the theory presented in our running example is not Convincing.

\subsection{Attempt}
As $\ileadsto \subseteq I × \allhist × \Phi$, the set of a priori possible $\ileadsto$ is $\powerset{I × \allhist × \Phi}$.

A claim $C$ is a subset of $\powerset{I × \allhist × \Phi}$.

\begin{theorem}
	\label{th:svalid}
	A normatively adequate empirical theory that has no falsification sequence is valid.
\end{theorem}

\subsection{Effectively comparing models}
\label{sec:compare}
Recall from \cref{sec:coh} that $\incompat = \set{(\prop, \prop') \suchthat \forall \hist \in \allhist: \hist \ileadsto \prop ⇒ \hist \nileadsto \prop'}$ denotes the set of pairs of incompatible propositions.
Given two empirical theories $\gamma$ and $\delta$ for a given normative theory, say that they disagree about $(\prop, \propbar) \in \incompat$ iff $\prop \in \gpropse[\gamma] \land \propbar \in \gpropse[\delta]$. Say that the theories are incompatible iff there is at least one pair $(\prop, \propbar) \in \incompat$ that they disagree about.

@@

>>>>

\subsection{Validity and falsification}
\label{sec:validity}
An empirical theory claims to be valid, in the following sense.
\begin{definition}[Validity]
	An empirical theory is \emph{valid} iff $\forall i \in I, \prop \in \gpropse: \prop \in \iprops$.
\end{definition}
To achieve this goal, it claims that it is able to produce decisive arguments for its supported propositions.
These arguments can vary according to the individual $i \in I$.
Given an individual, the theory claims that a decisive argument will be systematically found by starting from an initial argument given by $\gleadsto$, and following a chain of defense defined by $\dg$ as required, depending on the paths that challenge the arguments produced along the way, as judged by the individual $i$. The property that a theory manages to systematically produce decisive arguments in such a way is named Convincingness. Defining it precisely requires to define first the notion of required arguments.

@@

It is now necessary to verify that the claim of convincingness is, first, falsifiable, and second, sufficient to yield validity.

Let us first clarify what falsifiability formally means in our context. So far, convincingness is defined only for the “real” relation $\leadsto$; we now need it defined more generally. Given any relation $R \subseteq I × \allhist × \Phi$, say that a descriptive theory is convincing under $R$ whenever it is convincing, as defined above, but replacing $\leadsto$ by $R$. 
We can now speak counterfactually of the theory being convincing, or not, assuming the observable relation is $R$ instead of $\leadsto$. 
Observations of the relation $\leadsto$ can in general include positive statements, of the form $\hist \ileadsto \phi$, and negative ones, of the form $\hist \nileadsto \phi$. (One way to obtain such negative observations is to use an incompatible proposition $\phi'$ and deduce from $\hist \ileadsto \phi'$ that $\hist \nileadsto \phi$.) A relation $R \subseteq I × \allhist × \Phi$ is compatible with the positive observations $O \subseteq I × \allhist × \Phi$ and the negative observations $O' \subseteq I × \allhist × \Phi$ iff $O \subseteq R \land R \cap O' = \emptyset$.
The claim of convincingness is said to be falsifiable iff there exists finite set of observations $O, O' \subseteq I × \allhist × \Phi$ such that $\forall R \subseteq I × \allhist × \Phi \suchthat O \supseteq R \land R \cap O' = \emptyset$, the theory is not convincing under $R$. 
In such a case, no possible relation $R$ compatible with the observations $O, O'$ make the theory convincing; in other words, the observations $O, O'$ prove that the theory is not convincing.
%\begin{example}[Two claims]
%	Given a normative theory and $\phi \in \Phi$, the claim $C_1 = \set{{\leadsto} \subseteq I × \allhist × \Phi \suchthat \forall i \in I, \exists \ar \in \allargs \suchthat \ar \ileadstost \phi}$ corresponds to claiming that for all individuals $i \in I$, some argument is $i$-decisive for $\phi$. If the set of arguments is infinite, $C_1$ is not falsifiable.  
%	Given $\ar \in \allargs$ and $\phi \in \Phi$, the set $C_2 = \set{{\leadsto} \subseteq I × \allhist × \Phi \suchthat \forall i \in I, \ar \ileadstost \phi}$ denotes the claim that all individuals will consider the argument $\ar$ as decisive for $\phi$. This claim is falsified by any observation such that $\hist \nileadsto \phi$ for some $\hist \in \allhist$ with $\ar \in \histend$, thus, $C_2$ is falsifiable.
%\end{example}

The following theorem states that the descriptive theories defined here are falsifiable, unless they state nothing about anybody’s \ac{DJ}. \Cref{sec:prooffals} proves it and describes the set of falsifiers of such theories.
This theorem implicitly assumes full observability, that is, it supposes that any positive or negative observations are in principle observable and thus available for falsification attempts. Further works should investigate to which extent this hypothesis can be relaxed.
\begin{theorem}
	\label{th:convfals}
	The claim of convincingness is falsifiable iff ${\gleadsto} ≠ \emptyset$.
\end{theorem}
It can also be seen that if a theory is not constrained in its depth, its claim of convincingness may be unfalsifiable, thus the hypothesis of finite depth is required to obtain this theorem. The formal proof of this fact is omitted because it is not especially enlightening; that this fact holds can be seen by examining the use of this hypothesis in the proof of \cref{th:convfals}.

Only remains now to prove that convincingness, together with normative adequacy, is sufficient to obtain validity. The proof is also provided in \cref{sec:prooffals}.
\begin{theorem}
	\label{th:convvalid}
	A normatively adequate empirical theory that is convincing is valid.
\end{theorem}

\subsection{Effectively comparing models}
\label{sec:compare}
Recall from \cref{sec:coh} that $\incompat = \set{(\prop, \prop') \suchthat \forall \hist \in \allhist: \hist \ileadsto \prop ⇒ \hist \nileadsto \prop'}$ denotes the set of pairs of incompatible propositions.
\commentOC{Does this mandate completeness?}
Given two empirical theories $\gamma$ and $\delta$ for a given normative theory, say that they disagree about $(\prop, \propbar) \in \incompat$ iff $\prop \in \gpropse[\gamma] \land \propbar \in \gpropse[\delta]$. Say that the theories are incompatible iff there is at least one pair $(\prop, \propbar) \in \incompat$ that they disagree about.

@@

\section{Empirical validity, first}
\label{sec:prooffals}
To verify that the property of Convincingness is purely empirical, it is necessary to show that whenever a theory fails it, it is possible to see it using a set of observations of $\allleadsto$. Conversely, it is also desirable to ensure that only the observations that ensure that Convincingness is false count for falsifying a theory.
In other words, it is desirable to capture exactly which sets of observations falsify the claim of convincingness. This will also provide a proof of \cref{th:convfals}.

The following example may help to form an intuition about what the possible falsifying sets of observations are.
\begin{example}[Falsification sequences]
	Let us examine the possible observations that would suffice to show that the theory presented in our running example is not Convincing.
